---
main_topsize: 0.2 #percent coverage of the poster
main_bottomsize: 0.1
#ESSENTIALS
title: '**What Impacts World Happiness Scores?**'
author:
  - name: '**Elisabeth Foster**'
    affil: 1
    main: true
    email: ef5062a@american.edu
affiliation:
  - num: 1
    address: American University, School of International Service
main_findings:
  - "**Freedom** to make life choices impacts **World Happiness Scores**."
output: 
  posterdown::posterdown_betterport:
    self_contained: true
    pandoc_args: --mathjax
    number_sections: false
link-citations: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 

# setup 
setwd("/Users/elisabethfoster/Library/CloudStorage/OneDrive-Personal/American University Graduate School/Spring 2023/Data Analytics/R Work/Poster")

library(tidyverse)
library(ggplot2)
library(readxl)

happy <- read_excel("WHR2023.xls")


```

# Introduction
**Research Question:** Does freedom to make life choices effect happiness?  

**Expectations:** 

*Null hypothesis:* There is not a relationship between world happiness scores and freedom to make life choices. 

*Alternative hypothesis:* There is an effect of freedom to make life choices on world happiness scores. 

**Importance:** If there is a relationship between freedom to make life choices and world happiness scores, taking steps to increase a country's freedom score could be leveraged to improve their world happiness score. 


# Data

**Source:** This data comes from the [2023 World Happiness Project report](https://happiness-report.s3.amazonaws.com/2023/WHR+23_Statistical_Appendix.pdf). The majority of the data from their dataset comes from the Gallup World Poll, including the variables happiness score, social support, freedom to make life choices, generosity, and corruption perception. Some variables’ data came from other sources, including GDP per capita (World Bank’s World Development Indicators) and healthy life expectancy (World Health Organization’s Global Health Observatory data repository). 


```{r graph, ECHO = FALSE}
# graph for data section
happy1 = happy %>%
  rename("Happiness Score" = "Ladder score") %>%
  select(c("Country name", "Happiness Score", "Logged GDP per capita",
           "Social support", "Healthy life expectancy", 
           "Freedom to make life choices", "Generosity", 
           "Perceptions of corruption"))

ggplot(happy1, aes(x = `Happiness Score`)) +
  geom_histogram(color = "black",
                 fill = "#076060",
                 bins = 20) + 
  geom_vline(xintercept = mean(happy1$`Happiness Score`), 
             color = "black", linetype = "dashed") +
  theme_minimal() + 
  labs(
      title = bquote(~bold("World Happiness Scores 2023")),
      caption = "Source: World Happiness Report 2023",
      x = "Happiness Score",
      y = "Number of Countries"
  )
```
**Figure 1:** *Figure 1* demonstrates a count of the number of countries at each world happiness score in 2023. The dotted line represents the mean world happiness score. 

The world happiness score by country count distribution is slightly left skewed. The average happiness score is 5.5 and the median is 5.7. Afghanistan has the lowest happiness score at 1.9; it is one of two low level outliers, the other being Lebanon at 2.4. Finland has the highest score at 7.8. The highest score is not an outlier, and it is accompanied by multiple other countries with 7.0+ scores like Denmark and Iceland. 


```{r, ECHO = FALSE, fig.cap = 'Caption here.'}
# summary statistics table for data section 
library(vtable)
st(happy1, vars = c("Happiness Score", "Freedom to make life choices",
                    "Logged GDP per capita", "Perceptions of corruption"),
   col.width=NA)
```
**Table 1:** The table above demonstrates summary statistics for the variables that were considered when evaluating the relationship between world happiness score and freedom to make life choices. 


# Analysis
```{r, ECHO = FALSE}
ggplot(happy1, aes(x = `Happiness Score`, y = `Freedom to make life choices`)) +
  geom_point(color = "#076060",
                 bins = 20) + 
  theme_minimal() + 
  labs(
    title = bquote(~bold("World Happiness Scores and Freedom to make life choices")),
    caption = "Source: World Happiness Report 2023",
    x = "Happiness Score",
    y = "Freedom to make life choices"
  ) +
  geom_smooth(se=FALSE, color = "black")

```
**Figure 2:** This graph demonstrates the relationship between world happiness scores and freedom to make life choices, highlighted in greater depth in *Table 2*.


```{r, ECHO = FALSE, include = FALSE}
# linear regression 
mod = lm(`Happiness Score` ~ `Freedom to make life choices`, data = happy1)
mod

mod1 = lm(`Happiness Score` ~ `Freedom to make life choices` + 
            `Logged GDP per capita`, happy1)
mod1

mod2 = lm(`Happiness Score` ~ `Freedom to make life choices` +
          `Logged GDP per capita` + `Perceptions of corruption`, happy1)
```

```{r, results = 'asis', ECHO = FALSE}
# table
stargazer::stargazer(mod, mod1, mod2, type = "html", keep.stat = 'n')

```
**Table 2:** *Table 2* is a multivariate regression table that tests to see if there is a relationship between world happiness score and freedom to make life choices. Logged GDP per capita and perceptions of corruption were tested as potential confounding variables. 

Model 1 of *Table 2* demonstrates that freedom to make life choices has a positive and statistically significant relationship with world happiness scores. Models 2 and 3 show that this relationship holds even when potential confounding variables are introduced. For example, in Model 3, for every one unit increase in freedom to make life choices, there is a 3.759 unit increase in happiness score, holding logged GDP per capita and perceptions of corruption constant. 


# Findings
We can conclude that the freedom to make life choices variable independently has a positive, statistically significant relationship with the world happiness scores variable. For every one unit increase in the freedom to make life choices index, there is nearly a 4 unit increase in happiness score, holding logged GDP per capita and perceptions of corruption constant. 





